### 2024-05-22 README.Rmd
Still trying to pull a bunch of separate strands to obtain a coherent narrative

@hie_how_2023

@power_3_2024

It seemed inevitable that universities _had to do something_ in the face of all this innovative and perhaps only slightly scary technology.

In response to the rapid, and disruptive, release of tools that were announced in revolutionary terms, McMaster developed a set of provisional [guidelines](https://provost.mcmaster.ca/office-of-the-provost-2/generative-artificial-intelligence-2/generative-ai-for-staff/provisional-guidelines-on-the-use-of-generative-ai-in-operational-excellence/) (when where they developed and released? It is possible that this was done as a relatively urgent matter) to inform the university community about the use of AI in "operational excellence". These guidelines where released under a Creative Commons Licence and have since inspired those of other post-secondary institutions in Canada, including Trent University [-@trent_university_generative_2024] and the University of Guelph [-@office_of_teaching_and_learning_provisional_2023].

There is now an effort underway to update the preliminary guidelines (to make them permanent?) and McMaster, via the Office of the Provost, [launched a committee](#ai-committee-announced) to explore the use of AI across campus. This committee is composed of three sub-committees that examine the use of AI in 1) teaching and learning, 2) research, and 3) operational excellence in "our work" (the Provost's words)(probably meaning the university's work; not sure if the Provost themself uses AI).

@@

De-duplication of training data improves LLMs [@lee2022deduplicating]

No wonder: it amounts to having more different information.

@@

### Use of materials on the web

@@

@@

The example above may seem trivial, even amusing in the case above.It was not so funny for

@@

## As an aside

On antropomorphizing GenAI tools. This happens mostly in ways that are barely liminal. For example, in the document "Overview of AI at McMaster" we read that GenAI's somehow have an "ability to engage in nuanced conversation". This phrasing could be simply a literary device, but even so is indicative of a tendency to ascribe human attributes to a statistical model. Doing this carelessly risks amplifying existing biases. But in other cases, the effort is more blatant. Wrapping a statistical model in a human-like skin may actually be more devious, as in the new, sexy ChatGPT interface:

@@

Why? The reputational risk...universities are under siege from unsympathetic or actively hostile actors. Universities main source of power is their reputation as places that preserve and expand knowledge in a principled way. "This university uses ChatGPT" is a way to say that the university is redundant...an organization as expensive as an instution of higher-education is not needed to create spam and misinformation.

>>>>

@power_3_2024,^[Rhett Power is a motivational speaker and executive coach] for example, presents three innovative use cases for AI in higher education. First, AI can take over manual tasks and improve the budget by releasing administrative staff. AI can evaluate applications for admissions, including traits like the "grit" and "empathy" of applicants. AI can personalize the experience of consumers (also called students) by predicting their final grade with 60-70% accuracy, and by providing corrective and instructive feedback.
<!--release them in a very literal sense, as one advantage is to do _more_ with limited resources (which in the current fiscal environment means laying people off). The cases discussed by Mr. Power are questionable: there are already non-AI systems that deploy paperwork; the kind of information needed to identify students at risk of dropping out is probably quite invasive, and therefore impractical or unethical to collect; I very much doubt that it is legal to use AI in hiring and admissions. --> Mr. Power's message to the higher-eduation sector? It is time for institutions to level up: jump on the bandwagon, fellows, don't let it let you behind!

@hie_how_2023 argue that AI will likely be the primary way to access information in the future. For this reason, institutions must equip students with the skills needed to use AI effectively, in particular to perfect the art of _prompt engineering_, that is, the ability to craft queries that prompt an AI to provide the most useful outputs. Using AIs may improve learning and challenge the critical thinking of students. Faculty will also benefit from AI: it will be used to evaluate their students<!-- is this even legal? -->, and will handle repetitive tasks like writing course syllabi. @hie_how_2023^[Anthony Hié is Chief Innovation and Digital officer at [Excelia](https://www.excelia-group.com/about-excelia/our-schools/excelia-business-school). Writing in 2022, Mr. Hié predicted that the metaverse would be a [revolution in higher education](https://www.linkedin.com/pulse/metaverse-dedicated-pedagogy-revolution-higher-education-anthony-hi%C3%A9). In October 2023 facebook [slashed its metaverse team](https://www.reuters.com/technology/meta-lay-off-employees-metaverse-silicon-unit-wednesday-2023-10-03/). Microsoft [disbanded](https://www.theinformation.com/articles/microsoft-kills-its-industrial-metaverse-team-after-4-months) its Industrial Metaverse Core Team four months after launching it. Claire Thouary is the founder of [QACE UP](http://qace-up.fr), a consulting firm that supports higher-education institutions in their quest for "quality excellence".] advice that "AI is here to stay", that "[faculty] might not be able to stop the AI trend from growing", so they might as well get on with the program and try to shape it as best they can.

It seemed inevitable that universities _had to do something_ in the face of all this unavoidable, innovative technology that many feared might leave them behind.

The fear of missing out, combined with a modicum of concern for the ethical implications, has led to institutional responses that have been less than coherent. An example is the case of Emory University, where the university gave a $10,000 prize to a group of students participating in an Entrepreneurship Summit, only for the the students to be suspended later by the school's Honor Council for the potential of their tool for cheating^[There was no evidence that cheating had happened.] [@koebler__university_2024].

For its part, McMaster was a leader in responding to the rapid, and disruptive, release of _revolutionary_ tools, and its provisional [guidelines](https://provost.mcmaster.ca/office-of-the-provost-2/generative-artificial-intelligence-2/generative-ai-for-staff/provisional-guidelines-on-the-use-of-generative-ai-in-operational-excellence/) (released under a Creative Commons Licence) have since inspired those of other post-secondary institutions in Canada, including Trent University [-@trent_university_generative_2024] and the University of Guelph [-@office_of_teaching_and_learning_provisional_2023].

The aim of the provisional guidelines was to inform the university community about the use of AI in "operational excellence". But the time has come to update them, and the institution, via its Office of the Provost, [launched a committee](#ai-committee-announced) to explore the use of AI across campus. This committee is composed of three sub-committees that examine the use of AI in 1) teaching and learning, 2) research, and 3) operational excellence in "our work" (the Provost's words)(probably meaning the university's work; not sure if the Provost themself uses AI).

@@

De-duplication of training data improves LLMs [@lee2022deduplicating]. No wonder: it amounts to having more different information.

@@

### Extractive

@@

@germain_google_2023
@norem_intels_2023

### Exploitative

Above, the ways in which GenAI is exploitative of people all over the world, but particularly in the Global South were highlighted. But AI companies do not spare the rich either. OpenAI's Sam Altman, a self-avowed admirer of the movie _Her_, featuring Scarlet Johansson as an AI voice assistant, reportedly tried to hire Ms. Johannson to be the voice of ChatGPT 4.0 [@kastrenakes_scarlett_2024]. After Ms. Johannson decline, OpenAI went and used a mimicry of [her voice](@as-an-aside) anyways, which was retired after Ms. Johansson threatened legal action^[Altman's denials that OpenAI had not used Ms. Johansson's voice sounded hollow, after he [posted](https://x.com/sama/status/1790075827666796666) on social media the word "her" at the time the new version of ChatGPT was announced.] [@allyn_scarlett_2024]. Alas, very few have the social standing of Scarlett Johansson to challenge OpenAI. Most people in the world lack that power.

@williams_exploited_2022 (Summary needed)

@perrigo_2_2023 (Follows summary)

ChatGPT was trained on vast amounts of data scooped from the internet, where many societal problems are amplified. Trained on horrible content, it was hard to sell ChatGPT when it was prone to racist, sexist, obscene outbursts. But companies like facebook, who did not want to pay for moderation, had hit on a solution of sorts <!--: train an AI to detect content that was unpalatable.-->

The solution?

> [F]eed an AI with labeled examples of violence, hate speech, and sexual abuse, and that tool could learn to detect those forms of toxicity in the wild. That detector would be built into ChatGPT to check whether it was echoing the toxicity of its training data, and filter it out before it ever reached the user. It could also help scrub toxic text from the training datasets of future AI models.

The key here is "labeled examples". What are those labels? Many machine algorithms do not learn by themselves...they learn from humans, and need what is called "labelled" data: a images that humans have labelled as "dog", "cats", etc. Or bits of text that have been labelled as...well...questionable. Labelling data is extremely time consuming and labor intensive. What's a corporation to do?

> To get those labels, OpenAI sent tens of thousands of snippets of text to an outsourcing firm in Kenya, beginning in November 2021. Much of that text appeared to have been pulled from the darkest recesses of the internet. Some of it described situations in graphic detail like child sexual abuse, bestiality, murder, suicide, torture, self harm, and incest. "

What did this mean to the people doing all this labeling in one of the poorest regions in the world? First of all, an income: "data labelers employed...on behalf of OpenAI were paid a take-home wage of between around $1.32 and $2 per hour depending on seniority and performance." But at what cost?

> One Sama worker tasked with reading and labeling text for OpenAI told TIME he suffered from recurring visions after reading a graphic description of a man having sex with a dog in the presence of a young child. “That was torture,” he said. “You will read a number of statements like that all through the week. By the time it gets to Friday, you are disturbed from thinking through that picture."

@hao_hidden_2023

Quotes from Hao's podcast:

> Karen Hao: One of the reasons why ChatGPT was able to become so virally popular and continued to sustain popularity is because it is largely not spewing really awful things. People feel comfortable using the product knowing that it's not going to do that.

> Annie Minoff: At least it won't do it in English. If Alex wanted to use ChatGPT in his native language, Swahili, would he be able to do that?

> Karen Hao: You can interact with ChatGPT in Swahili, but ChatGPT was developed primarily to work in English. So a lot of the scrubbing, the content moderation, the important safety measures within the chatbot were done in English. So when you prompt it in Swahili, you'll get more misinformation. You'll get more confusing sentences that don't make sense, and you will potentially get more of this content that they worked so hard to filter out because they were only filtering it in English.

Think about linguistic hegemony, and how the models are only as good (or as bad) as their inputs.

@@

The example above may seem like an amusement, but a GenAI is not a toy, as Air Canada discover when the company was found liable for "negligent misrepresentation", after a court denied claims by the airline that the chatbot was somehow responsible for its own actions as an effectively separate legal entity [@garcia_what_2024].

Similar GenAI-generated mishaps abound.

@@

## As an aside {#as-an-aside}

On antropomorphizing GenAI tools. This happens mostly in ways that are barely liminal. For example, in the document "Overview of AI at McMaster" we read that GenAI's somehow have an "ability to engage in nuanced conversation". This phrasing could be simply a literary device, but even so it is indicative of a tendency to ascribe human attributes to a statistical model. Doing this risks amplifying existing biases. But in other cases, the effort is more blatant. Wrapping a statistical model in a human-like skin may actually be more devious, as in the new, sexy ChatGPT interface:

@@

Why? The reputational risk...universities are under siege from unsympathetic or actively hostile actors. Universities main source of power is their reputation as places that preserve and expand knowledge in a principled way. "This university uses ChatGPT" is a way to say that the university is redundant...an organization as expensive as an institution of higher-education is not needed when there are cheaper alternatives for such an education ("act now before you are laid off!"):

```{r ligemy-ai-bundle, out.width="50%", fig.cap="\\label{fig:ai-success-bundle}An AI success bundle from the \"The Real People of Ligency PR and Marketing Team\""}
bundle_plot <- ggplot() +
  labs(caption = "https://www.udemy.com/user/ligency-team/") +
  theme_void()

ggdraw() + 
  draw_plot(bundle_plot) +
  draw_image(image = "images/trail-blazing-ai-bundle.png")
```
